Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-07-06 paper/submitted/round2_png/reviewer_response.tex

\begin{document}

%%%%% ASK KAREN/ELIZABETH
%do we need to refer to line numbers?
%if a comment asks for clarification, do we reproduce it here or point to the changed lines/sections?

\section*{Response to Referee Comments}

We thank the reviewers for their detailed and insightful comments and helping us strengthen the manuscript. Our point-by-point responses to the critique are outlined below, with reviewer comments in black and \blue{responses in blue}.

\answer{Content was rearranged.}

\answer{We added text about the use cases to the second paragraph of the introduction, at the bottom of page 2.}

\answer{We corrected the reference.}

\answer{We corrected the reference.}

\answer{We added a sentence to the final paragraph to this effect.}

\answer{We elaborated upon the similarities with, and differences from, [7] in the last two paragraphs of Section 2.1.}

\answer{We have made major revisions to Section 2. We moved the description of the low- and high-fidelity models earlier, to Section 2.1, and elaborated upon them. We expanded our explanation of compatibility at the end of Section 3.1. References to Propositions and proofs in [7] have been replicated within the manuscript and adapted to our setting. We added an additional proposition as suggested.}

\answer{We moved the introduction of the first and simplest numerical example to Section 3.1 and added a description of Algorithm 1 in terms of this example.}
%egh, does the new explanation make sense?

\answer{References to Propositions and proofs in [7] have been replicated within the manuscript and adapted to our setting.}

\answer{We reworded the description of the parameters in Section 4.2.1 to be more clear/explicit about the relevant spaces and representations. The mesh was sufficiently refined that further refinements did not significantly change the solutions.}

It could be illustrative to see how the parameters $q$ change with the refinement process.

\answer{Figure 3 is intended to show the change in QoI error with refinement, until the target QoI is reached. We modified Figure 3 to be on log scale and included a line denoting the target QoI error. We removed the 100\% refinement point, since it did not actually correspond to a refinement iteration.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 4}

The paper is basically an application of [7] to modeling error demonstrated with rather transparent problems in 2D flow. The authors ignore the first papers on Goal-Oriented Estimation and modeling error laid out in the papers of Oden and Vemaganti in JCP in 2000, and extended by these authors in CMAME in 2001. The relationship between the current work and these papers should be clearly explained and cited. These revisions and re-review would be advised.

\answer{We apologize for not referencing the aforementioned papers; we missed them in our literature review and did not intend to imply their irrelevance. Citations of these papers have been added to the fourth paragraph of the introduction.}

### 2017-07-06 paper/submitted/round2_png/algorithms_v4_cmame.tex
edits to paper, after karen's comments

As an example, consider a pair of low- and high-fidelity models described by the convection-diffusion and convection-diffusion-reaction equations
\begin{subequations}
\begin{align}
\textrm{low-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u = q, \label{eq:cdvcdrLF}\\
\textrm{high-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2 = q, \label{eq:cdvcdrHF}
\end{align}
\end{subequations}
where $k_d$ and $k_r$ are diffusion and reaction coefficients, respectively, and the state $u$ and parameters $q$ are continuous functions over a domain $\Omega$ with homogenous Dirichlet boundary conditions. To form the mixed-fidelity models, we divide the domain into complementary subdomains, $\Omega_{HF}$ and $\Omega_{LF}$, where the high- and low-fidelity models are solved, respectively; initially, $\Omega_{LF}=\Omega$. The resulting mixed-fidelity models can be described by,
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k^{MF}_ru^2= q,
\end{equation}
%
where $k^{MF}_r$ is a piecewise-constant reaction coefficient, %that's zero in the $\Omega_{LF}$ and equal to $k_r$ in $\Omega_{HF}$; 
%
\begin{equation}
k^{MF}_r=
\begin{cases}
k_r & \textrm{if }x\in\Omega_{HF} \\
0 & \textrm{if }x\in\Omega_{LF}.
\end{cases}
\end{equation}
%
Let the QoI be described by $I(q,u)=\int_{\Omega_I} u \:\textrm{d}\Omega$, where $\Omega_I\subset\Omega$ is the QoI region. Let the observations be of the state at some set of locations, so that the data-misfit term is $\frac{1}{2}\int_\Omega \delta_{obs}(u-d)^2\:\textrm{d}\Omega$, where $\delta_{obs}$ is a sum of Dirac delta functions indicating the locations of the observations. Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}, with $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla q\|_2^2\:\textrm{d}\Omega$, where $\beta$ is a regularization coefficient. Assuming divergence-free velocity $\vec{V}$, the mixed-fidelity model has augmented Lagrangian
\begin{eqnarray}
\mathcal{M}_{MF}((q,u,z)(p,v,y)) &=& \int_{\Omega_I\subset\Omega} u \:\textrm{d}\Omega \nonumber\\
&+& \int_\Omega p(-\beta\nabla^2q+z) \nonumber\\
&+& v(k_d\nabla^2z+\vec{V}\cdot\nabla z+2k^{MF}_rzu+\delta_{obs}(u-d)) \nonumber\\ 
&+& y(k_d\nabla^2u-\vec{V}\cdot\nabla u+k^{MF}_ru^2+q)\:\textrm{d}\Omega.
\end{eqnarray}
Taking variations with respect to the auxiliary variables $(p,v,y)$ recovers the equations for the primary variables; taking variations with respect to the primary variables gives the equations for the auxiliary variables. Solving for the primary variables and then the auxiliary variables gives the stationary point $\Psi_{MF}$ of $\mathcal{M}_{MF}$. The weighted residual of the high-fidelity primary and auxiliary equations, evaluated at $\Psi_{MF}$, describes a linear functional in the weights (see \Cref{eq:supadjout}). Taking the adjoint of these high-fidelity primary and auxiliary systems, with output described by this linear functional, gives the supplementary adjoint; the components of the supplementary adjoint corresponding to the primary and auxiliary systems can be solved separately. Having obtained $\Psi_{MF}$ and the supplementary adjoint $\Lambda$, we evaluate the error estimate (see Step \ref{alg-line:MFerrorest} of \Cref{alg:refSeries}), localize it (see \Cref{sec:errLocal}), and use the localization to guide the selection of elements currently in $\Omega_{LF}$ to include in $\Omega_{HF}$ in the next iteration. We return to this example and give results of applying \Cref{alg:refSeries} to this pair of models in \Cref{sec:cdvcdr}.

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, be a reduced-order model, or have a reduced parameter and/or state space. The derived error estimate is not applicable to all models, however. It is assumed that the two models share the same domain; for example, the low-fidelity model can not be a one-dimensional average of a two-dimensional high-fidelity model. \red{The high-fidelity model must by expressible in a weak form}, so this cannot be applied to, for example, a high-fidelity model of chemical reactions using kinetic Monte Carlo. %if the LF model does not have a weak form, say it's from a response surface...what would the components of Psi be?
We need some degree of compatibility between the two models, in that we require the low-fidelity solutions to be interpretable in the context of the high-fidelity model; specifically, we require that $\Psi_{LF}$ be in a space admissible to $\mathcal{M}'_{HF}$ and $\mathcal{M}_{HF}$. For example, compared to the high-fidelity model, low-fidelity model may have a coarser mesh, basis functions of a lower order, or basis functions spanning a reduced subspace. The high-fidelity model could describe changes in states that have a fixed nominal value in the low-fidelity model; the high-fidelity model could allow anisotropy of the inferred parameters while the low-fidelity model could constrain their components and require isotropy. We note that in the case where the two models correspond to two levels of mesh refinement, method described in~\cite{BecVex05} would likely be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead.
%, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$. %implied by admissibility of LF to HF

>>>>

As an example, consider a pair of low- and high-fidelity models described by the convection-diffusion and convection-diffusion-reaction equations
\begin{subequations}
\begin{align}
\textrm{low-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u = q, \label{eq:cdvcdrLF}\\
\textrm{high-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2 = q, \label{eq:cdvcdrHF}
\end{align}
\end{subequations}
where $k_d$ and $k_r$ are diffusion and reaction coefficients, respectively, and the state $u$ and parameters $q$ are continuous functions over a domain $\Omega$ with homogenous Dirichlet boundary conditions. To form the mixed-fidelity models, we divide the domain into complementary subdomains, $\Omega_{HF}$ and $\Omega_{LF}$, where the high- and low-fidelity models are solved, respectively; initially, $\Omega_{LF}=\Omega$. The resulting mixed-fidelity models can be described by,
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k^{MF}_ru^2= q,
\end{equation}
%
where $k^{MF}_r$ is a piecewise-constant reaction coefficient, %that's zero in the $\Omega_{LF}$ and equal to $k_r$ in $\Omega_{HF}$; 
%
\begin{equation}
k^{MF}_r=
\begin{cases}
k_r & \textrm{if }x\in\Omega_{HF} \\
0 & \textrm{if }x\in\Omega_{LF}.
\end{cases}
\end{equation}
%
Let the QoI be described by $I(q,u)=\int_{\Omega_I} u \:\textrm{d}\Omega$, where $\Omega_I\subset\Omega$ is the QoI region. Let the observations be of the state at some set of locations, so that the data-misfit term is $\frac{1}{2}\int_\Omega \delta_{obs}(u-d)^2\:\textrm{d}\Omega$, where $\delta_{obs}$ is a sum of Dirac delta functions indicating the locations of the observations. Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}, with $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla q\|_2^2\:\textrm{d}\Omega$, where $\beta$ is a regularization coefficient. Assuming divergence-free velocity $\vec{V}$, the mixed-fidelity model has augmented Lagrangian
\begin{eqnarray}
\mathcal{M}_{MF}((q,u,z)(p,v,y)) &=& \int_{\Omega_I\subset\Omega} u \:\textrm{d}\Omega \nonumber\\
&+& \int_\Omega p(-\beta\nabla^2q+z) \nonumber\\
&+& v(k_d\nabla^2z+\vec{V}\cdot\nabla z+2k^{MF}_rzu+\delta_{obs}(u-d)) \nonumber\\ 
&+& y(k_d\nabla^2u-\vec{V}\cdot\nabla u+k^{MF}_ru^2+q)\:\textrm{d}\Omega.
\end{eqnarray}
Taking variations with respect to the auxiliary variables $(p,v,y)$ recovers the equations for the primary variables; taking variations with respect to the primary variables gives the equations for the auxiliary variables. Solving for the primary variables and then the auxiliary variables gives the stationary point $\Psi_{MF}$ of $\mathcal{M}_{MF}$. 
The high-fidelity versions of these primary and auxiliary systems are satisfied by $\Psi_{HF}$; we seek the error in the linear functional described by \Cref{eq:supadjout} induced by applying it to $\Psi_{MF}$ instead of $\Psi_{HF}$. We take the classical dual-weighted residual approach and solve for the supplementary adjoint $\Lambda$, linearizing about $\Psi_{MF}$ instead of $\Psi_{HF}$ when necessary, since the latter is not available; the components of the supplementary adjoint corresponding to the primary and auxiliary systems can be solved separately. Having obtained $\Psi_{MF}$ and the supplementary adjoint $\Lambda$, we evaluate the error estimate (see Step \ref{alg-line:MFerrorest} of \Cref{alg:refSeries}), localize it (see \Cref{sec:errLocal}), and use the localization to guide the selection of elements currently in $\Omega_{LF}$ to include in $\Omega_{HF}$ in the next iteration. We return to this example and give results of applying \Cref{alg:refSeries} to this pair of models in \Cref{sec:cdvcdr}.
%The weighted residual of the high-fidelity primary and auxiliary equations, evaluated at $\Psi_{MF}$, describes a linear functional in the weights (see \Cref{eq:supadjout}). Taking the adjoint of these high-fidelity primary and auxiliary systems, with output described by this linear functional, gives the supplementary adjoint; the components of the supplementary adjoint corresponding to the primary and auxiliary systems can be solved separately. Having obtained $\Psi_{MF}$ and the supplementary adjoint $\Lambda$, we evaluate the error estimate (see Step \ref{alg-line:MFerrorest} of \Cref{alg:refSeries}), localize it (see \Cref{sec:errLocal}), and use the localization to guide the selection of elements currently in $\Omega_{LF}$ to include in $\Omega_{HF}$ in the next iteration. We return to this example and give results of applying \Cref{alg:refSeries} to this pair of models in \Cref{sec:cdvcdr}.

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena, or be a reduced-order model, or have a reduced parameter and/or state space. The derived error estimate is not applicable to all models, however. It is assumed that the two models share the same domain; for example, the low-fidelity model can not be a one-dimensional average of a two-dimensional high-fidelity model. The low- and high-fidelity models must be expressible in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. We need some degree of compatibility between the two models, in that we require the low-fidelity solutions to be interpretable in the context of the high-fidelity model; specifically, we require that $\Psi_{LF}$ be in a space admissible to $\mathcal{M}'_{HF}$ and $\mathcal{M}_{HF}$. For example, compared to the high-fidelity model, the low-fidelity model may have a coarser mesh, basis functions of a lower order, or basis functions spanning a reduced subspace. The high-fidelity model could describe changes in states that have a fixed nominal value in the low-fidelity model; the high-fidelity model could allow anisotropy of the inferred parameters while the low-fidelity model could constrain their components and require isotropy. We note that in the case where the two models correspond to two levels of mesh refinement, method described in~\cite{BecVex05} would likely be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead.
%, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$. %implied by admissibility of LF to HF
